Shaking out the idiosyncracies

### 2020-02-05 analysis/reports/vetting.Rmd
scaled histograms

ggplot(data = filter(all_di, nsamples >= 2000, singletons == FALSE), aes(x = skew_percentile)) +
  geom_histogram(bins = 100) +
  theme_bw()

ggplot(data = filter(all_di, nsamples >= 2000, singletons == FALSE), aes(x = simpson_percentile)) +
  geom_histogram(bins = 100) +
  theme_bw()

@@

@@

ggplot(data = all_di_macd, aes(x = skew_percentile)) +
  geom_histogram(bins = 100) +
  theme_bw()

ggplot(data = all_di_macd, aes(x = simpson_percentile)) +
  geom_histogram(bins = 100) +
  theme_bw()
```

@@

ggplot(data = all_di_macd, aes(x = skew_percentile)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(treatment), scales = "free_y")

ggplot(data = all_di_macd, aes(x = simpson_percentile)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(treatment), scales = "free_y")

@@

** I AM REALLY NOT CONFIDENT IN THE DATA HANDLING HERE, NEED TO REVISIT WHEN SHARPER **

```{r macd ctrlcomp}

macd_comparisons <- read.csv(here::here("working-data", "macdb_data", "orderedcomparisons.csv"), header = F, stringsAsFactors = F)

colnames(macd_comparisons) <- c("studyID", "control", "site")
macd_comparisons <- macd_comparisons %>%
  mutate(site = as.character(site), control = as.character(control))

cc_di <- all_di_macd %>%
  filter(singletons == FALSE, treatment == "comparison") %>%
  left_join(macd_comparisons, by = c("studyID", "site")) %>%
  select(dat, site, skew_percentile, simpson_percentile, studyID, control) %>%
  rename(comparison = site) %>%
 left_join(select(all_di_macd, skew_percentile, simpson_percentile, site, singletons, nsamples), by = c("control" = "site")) %>%
  distinct() %>%
  rename(comp_skew = skew_percentile.x, comp_simp = simpson_percentile.x,
         control_skew = skew_percentile.y, control_simp = simpson_percentile.y)

ggplot(data = cc_di, aes(x = comp_skew, y = control_skew)) +
  geom_point(alpha = .5) +
#  xlim(0, 100) +
 # ylim(0, 100) +
  theme_bw() +
  geom_abline(intercept = 0, slope = 1, color = "green")

ggplot(data = cc_di, aes(x = comp_simp, y = control_simp)) +
  geom_point(alpha = .5) +
 xlim(0, 100) +
 ylim(0, 100) +
  theme_bw() +
  geom_abline(intercept = 0, slope = 1, color = "green")

@@

ggplot(data = all_di_p, aes(x = skew_percentile)) +
  geom_histogram(bins = 100) +
  theme_bw()

ggplot(data = all_di_p, aes(x = simpson_percentile)) +
  geom_histogram(bins = 100) +
  theme_bw()

@@

```{r portal trtmt}
ggplot(data = filter(all_di_p, singletons == F), aes(x = skew_percentile)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(season, treatment), scales = "free_y")

@@

ggplot(data = filter(all_di_p, singletons == F), aes(x = simpson_percentile)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(season, treatment), scales = "free_y")

>>>>

ggplot(data = filter(all_di, nsamples >= 2000, singletons == FALSE), aes(x = skew_percentile)) +
  geom_histogram(bins = 100, aes(y = stat(count / sum(count)))) +
  theme_bw() +
  geom_hline(yintercept = .01, linetype = 3)

ggplot(data = filter(all_di, nsamples >= 2000, singletons == FALSE), aes(x = simpson_percentile)) +
  geom_histogram(bins = 100, aes(y = stat(count / sum(count)))) +
  theme_bw() +
    geom_hline(yintercept = .01, linetype = 3)

sum(filter(all_di, nsamples >= 2000, singletons == F)$skew_percentile >= 95) / nrow(filter(all_di, nsamples >= 2000, singletons == F))
```

@@

## S and N v percentiles

```{r sn percentiles}

ggplot(data = filter(all_di, singletons == FALSE), aes(x = log(s0), y = log(n0), color = skew_percentile)) +
  geom_point(alpha = .1) +
  theme_bw() +
  scale_color_viridis_c(option = "plasma", end = .8)

ggplot(data = filter(all_di, singletons == FALSE, nsamples >= 2000), aes(x = log(s0), y = log(n0), color = skew_percentile)) +
  geom_point(alpha = .25) +
  theme_bw() +
  scale_color_viridis_c(option = "plasma", end = .8)

ggplot(data = filter(all_di, singletons == FALSE), aes(x = log(s0), y = log(n0), color = simpson_percentile)) +
  geom_point(alpha = .1) +
  theme_bw() +
  scale_color_viridis_c(option = "plasma", end = .8, direction = -1)

ggplot(data = filter(all_di, singletons == FALSE, nsamples >= 2000), aes(x = log(s0), y = log(n0), color = simpson_percentile)) +
  geom_point(alpha = .25) +
  theme_bw() +
  scale_color_viridis_c(option = "plasma", end = .8, direction = -1)
```

Even when you get >2k samples, it looks to me like you get *less unusual* observed SADs in the relatively low N/S region, which is also the region I expect to be most even.

```{r avg n percentile}

ggplot(data = filter(all_di, nsamples >= 2000, singletons == F), aes(x = log(n0/s0), y = skew_percentile)) +
  geom_point(alpha = .1) +
  theme_bw()

ggplot(data = filter(all_di, nsamples >= 2000, singletons == F), aes(x = log(n0/s0), y = simpson_percentile)) +
  geom_point(alpha = .1) +
  theme_bw()
```

You really don't get non-extreme percentiles in the high n/s region.

These scatterplots are good for seeing where the variation is but not the density; there's a *lot* of points on top of each other down at 0/up at 100.

@@

ggplot(data = all_di_macd, aes(x = skew_percentile)) +
   geom_histogram(bins = 100, aes(y = stat(count / sum(count)))) +
  theme_bw() +
    geom_hline(yintercept = .01, linetype = 3)

ggplot(data = all_di_macd, aes(x = simpson_percentile)) +
  geom_histogram(bins = 100, aes(y = stat(count / sum(count)))) +
  theme_bw() +
    geom_hline(yintercept = .01, linetype = 3)
```

@@

ggplot(data = all_di_macd, aes(x = skew_percentile)) +
    geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y")

ggplot(data = all_di_macd, aes(x = simpson_percentile)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y")

@@

<!-- ** I AM REALLY NOT CONFIDENT IN THE DATA HANDLING HERE, NEED TO REVISIT WHEN SHARPER ** -->

<!-- ```{r macd ctrlcomp} -->

<!-- macd_comparisons <- read.csv(here::here("working-data", "macdb_data", "orderedcomparisons.csv"), header = F, stringsAsFactors = F) -->

<!-- colnames(macd_comparisons) <- c("studyID", "control", "site") -->
<!-- macd_comparisons <- macd_comparisons %>% -->
<!--   mutate(site = as.character(site), control = as.character(control)) -->

<!-- cc_di <- all_di_macd %>% -->
<!--   filter(singletons == FALSE, treatment == "comparison") %>% -->
<!--   left_join(macd_comparisons, by = c("studyID", "site")) %>% -->
<!--   select(dat, site, skew_percentile, simpson_percentile, studyID, control) %>% -->
<!--   rename(comparison = site) %>% -->
<!--  left_join(select(all_di_macd, skew_percentile, simpson_percentile, site, singletons, nsamples), by = c("control" = "site")) %>% -->
<!--   distinct() %>% -->
<!--   rename(comp_skew = skew_percentile.x, comp_simp = simpson_percentile.x, -->
<!--          control_skew = skew_percentile.y, control_simp = simpson_percentile.y) -->

<!-- ggplot(data = cc_di, aes(x = comp_skew, y = control_skew)) + -->
<!--   geom_point(alpha = .5) + -->
<!-- #  xlim(0, 100) + -->
<!--  # ylim(0, 100) + -->
<!--   theme_bw() + -->
<!--   geom_abline(intercept = 0, slope = 1, color = "green") -->

<!-- ggplot(data = cc_di, aes(x = comp_simp, y = control_simp)) + -->
<!--   geom_point(alpha = .5) + -->
<!--  xlim(0, 100) + -->
<!--  ylim(0, 100) + -->
<!--   theme_bw() + -->
<!--   geom_abline(intercept = 0, slope = 1, color = "green") -->

<!-- ``` -->

@@

ggplot(data = all_di_p, aes(x = skew_percentile)) +
  geom_histogram(bins = 100, aes(y = stat(count / sum(count)))) +
  theme_bw() +
    geom_hline(yintercept = .01, linetype = 3)

ggplot(data = all_di_p, aes(x = simpson_percentile)) +
   geom_histogram(bins = 100, aes(y = stat(count / sum(count)))) +
  theme_bw() +
    geom_hline(yintercept = .01, linetype = 3)

@@

```{r portal trtmt}
ggplot(data = filter(all_di_p, singletons == F), aes(x = skew_percentile)) +
    geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y")

@@

ggplot(data = filter(all_di_p, singletons == F), aes(x = simpson_percentile)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y")
